STAT 672 Final Project:\\ Stochastic Gradient Descent

### 2018-04-17 paper/paper.tex
svm

In theory, SGD is agnostic to $f$ and $w$ as long as they meet required
assumptions, and so can be used in many different areas of optimization. In
practice, SGD is overwhelmingly popular in the statistical learning
field for estimating weight coefficients in predictive models. 
This section lays out some reasons for SGD's popularity; conducts a case study
using SGD; and identifies notable real-world applications of SGD.

\subsection{SGD and Statistical Learning}
\cite{shalev2011pegasos}

Loosely following the set-up of , consider a typical supervised classification problem. We have
feature matrix $\bm{X}_{n \times D, n \in \mathbb{R}^n, D \in \mathbb{R}^D}$ (corresponding to $n$ observations and $D$
features) and labels $\bm{Y}_{n \times 1}, Y_i \in \{-1, 1\}$. The goal is to
predict a particular observation's label $Y_i$ using that observation's features $\bm{X}_i$. 
We have a hypothesis class $\mathcal{F}$ consisting of
various functions $f_{\bm{w}}(\bm{X}_i) \in \mathcal{F}$ parametrized by weight
vector $\bm{w}$. We have loss function $L(Y_i, f_{\bm{w}}(\bm{X}_i))$ that
expresses the cost of mis-classification. Assume for now that $L$ is convex. 
We will consider the optimal function $\hat{f}_{\bm{w}}(\bm{X}_i) \in F$ 
that which minimizes empirical risk over all observations: $\frac{1}{n}
\sum_{i=1}^n L(Y_i, f_{\bm{w}}(\bm{X}_i))$.
Denote $\hat{\bm{w}}$ the weight coefficients of this optimal function. We thus
have:

@@

Because $L$ is convex, there is one critical point,
it is the global minimum, and it is located where the gradient of the loss
function with respect to $\bm{w}$ is zero. GD is an iterative
algorithm that can be used to numerically approximate this point. We first state
it in a general form, and then adapt it to the specific problem outlined above.

\begin{equation}
	\bm{w}_{t+1} = \bm{w}_t - \gamma \frac{1}{n}\sum_{i=1}^n
	\nabla L(Y_i, f_{\bm{w}_t}(\bm{X}_i))
\end{equation}

$t$ refers to iterations. $\gamma$ is a parameter controlling the step
size (also called ``learning rate''). Assume for now that $\gamma$ is fixed. For
$t=0$ we must supply an initial starting guess, $\bm{w}_0$.

\begin{equation}
	\left|\frac{1}{n}\sum_{i=1}^n \nabla L(\bm{Y}_i,
	f_{\bm{w}_t}(\bm{X}_i))\right| \leq \epsilon
\end{equation}

Note that the batch GD algorithm presented in (2) requires evaluating the
gradient for every single observation $i$. In a large-$n$ dataset, this can be
computationally infeasible. SGD's innovation is to instead evaluate only a
single randomly chosen observation $i$ at each iteration $t$.

\begin{equation}
	\bm{w}_{t+1} = \bm{w}_t - \gamma 
	\nabla L(Y_i, f_{\bm{w}_t}(\bm{X}_i))
\end{equation}

>>>>

In theory, SGD is agnostic to $f$ and $w$ as long as they meet required
assumptions, and so can be used in many different areas of optimization. In
practice, SGD is overwhelmingly popular in the statistical learning
field for estimating weight coefficients in predictive models.

\subsection{SGD and Statistical Learning}

Consider a typical supervised classification problem. We have
feature matrix $\bm{X}_{n \times D}$ (corresponding to $n$ observations and $D$
features) and labels $\bm{Y}_{n \times 1}, Y_i \in \{-1, 1\}$. The goal is to
predict a particular observation's label $Y_i$ using that observation's features
$\bm{X}_i$. To emphasize the utility of SGD, suppose that both $n$ and $D$ are
very large.
We have a hypothesis class $\mathcal{F}$ consisting of
various functions $f_{\bm{w}}(\bm{X}_i) \in \mathcal{F}$ parametrized by weight
vector $\bm{w}$. We have convex loss function $L(Y_i, f_{\bm{w}}(\bm{X}_i))$ that
expresses the cost of misclassification. 
We will consider the optimal function $\hat{f}_{\bm{w}}(\bm{X}_i) \in F$ 
that which minimizes empirical risk over all observations: $\frac{1}{n}
\sum_{i=1}^n L(Y_i, f_{\bm{w}}(\bm{X}_i))$.
Denote $\hat{\bm{w}}$ the weight coefficients of this optimal function. We thus
have:

@@

How shall we solve this optimization problem? Newton's method requires
calculating the Hessian, which will have $D^2$ entries; since $D$ is large,
this is inconvenient to compute. We could try batch gradient descent:

\begin{equation}
	\bm{w}^{(t+1)} = \bm{w}^{(t)} - \gamma \frac{1}{n}\sum_{i=1}^n
	\nabla_{\bm{w}} L(Y_i, f_{\bm{w}^{(t)}}(\bm{X}_i))
\end{equation}

But, in line with our analysis in section 2.2.2, batch GD requires evaluating
the gradient over every observation $i \leq n$ for every iteration $t$. Since we have defined $n$ to be
large, this is computationally inconvenient. Rather, we use SGD to evaluate only
a single randomly chosen observation per iteration.

\begin{equation}
	\bm{w}^{(t+1)} = \bm{w}^{(t)} - \nabla_{\bm{w}} L(Y_i, f_{\bm{w}^{(t)}}(\bm{X}_i))
\end{equation}

Per previous analyses we know that this approach will converge to an arbitrarily
good answer (if certain assumptions are met); will converge to that arbitrary
degree of precision in less time than plain gradient descent; does not
necessarily require computing the Hessian; and can be parallelized across many
machines. These attractive qualities make SGD well-represented in almost all areas of
machine learning. In the following subsection, we take a closer look at one
particularly prominent applications, support vector machines.

\subsubsection{Support vector machines}

Originally proposed by \cite{boser1992training} and \cite{cortes1995support}, support vector machines (SVM)
are among the best supervised learning algorithms. Here we set up the SVM
problem and then show how SGD can be used to solve it.

\cite{shalev2011pegasos}
